
Today there are various real-world problems that are (tried to be) solved by computational methods. One of them which 
is currently attacked at the Konrad Zuse Zentrum für Informationstechnik Berlin (ZIB) is the Problem of transporting 
natural gas through a network of pipelines. Motivated by this challenging problem the question of how to produce 
acyclic network flows arises. This problem is not specific for gas networks but for general flow in networks with arc 
capacities and indeed can be seen as just a special case of the famous minimum cost flow problem. 

To my knowledge there is no literature about acyclic network flows despite wide research on various flow problems. 
Whereever we have a network with nonpositive lower capacity bounds, we can make any flow acyclic by shifting flow 
back on the cycles until an edge has zero flow. 
However, we are interested in bounds on flow values for the arcs of the network. For each arc we have a maximum 
capacity and want to know the maximum flow value this arc could have in a feasible network flow that is acyclic. Thus 
we have to find the \textit{best} acyclic flow for each arc. This problem turns out to be much more challenging.


At first glance it may seem that solving a normal Min-Cost-Flow algorithm could give an appropriate bound 
on an arc. If we set the weight $w_e$ on an arc $e$ to 1 or -1 and all other weights to zero, we get 
a valid bound on the maximum or minimum possible flow on the arc. However, this bound is for most of the arcs far from 
reality. The reason is that a Min-Cost-Flow could have cycles within it and the algorithm could maximize the flow by 
sending as much flow as possible on a cycle. In natural gas networks this can not happen (except from compressor 
stations) since the flow is always induced by pressure differences.\\

\subsection{Outline}
The intention of this thesis is to investigate the problem of flow bounds in a network where the flow is assumed to be 
acyclic. 

We will first define the mathematical/algorithmical problem and examine connections to known problems 
and relaxations of the acyclic flowbound problem. We show that the minimum cost flow problem with general (possibly 
negative) weights becomes $\mathcal{NP}$-hard if we demand acyclic flow. 
Justified by this we model the problem as a mixed integer program. We present two different mixed integer programs 
that model our exact problem. One of the MIP models relies on constraints which each forbid cyclic flow on a cycle of 
the network. A network can have an exponential number of cycles. We show a condition under which an acyclicity 
constraint becomes redundant. Depending on the network we are given this could reduce the number of these constraints 
to be linear in the input size. Still on some networks the condition is not applicable.
We furthermore discuss the idea of an approach that tries to modify and adjust path based algorithms for maximum flow 
or minimum cost flow in order to solve this problem. We present examples to show that this can only produce a 
heuristic algorithm or else requires to solve another hard problem as subroutine. A heuristic algorithm close to the 
one we would get there is described and analysed.

The algorithms/models we discussed were also implemented. In the end we present details of the algorithms that were 
implemented and compare results achieved by them. 

\input{motivation}


% [TODO kurzer Abriss pber den Inhalt der Arbeit]
% Today more and more real-world problems in the areas of simulation and optimization are solved by mathematical and 
% computational methods. A growing number of these problems can be solved without problems, i.e. even huge instances give 
% an optimal or near optimal solution within seconds. Still, there remain problems that even on modern computers are hard 
% to solve. For these problems it is important to find ways to increase the efficiency of the algorithms. 
% 
% The topic of this thesis arises from the computation of flow in natural gas networks, which is currently developed 
% in the FORNE Project in a cooperation of OGE with universities and research insitutes including ZIB.
% %TODO genaueres zu FORNE? genaueres zum aufbau des Gasnetzes?
% The flow of natural gas in a network is described by nonlinear equations and depends on many parameters, which makes 
% the problem hard to solve. If we can find good upper and lower bounds for the flow on an arc during the preprocessing, 
% we can hope to improve the behavior of the nonlinear solver by giving these tigther bounds. 
% 
% The flow is induced by pressure differences, so in reality there can't be cyclic flow (if we exclude compressor 
% stations). Without the condition of acyclic flow, it is sufficient to run a standard min-cost-flow algorithm where the 
% maximized arc $e$ gets weight $w_e = -1$ and all others are 0. However, the arising bounds are far from optimal. If arc 
% $e$ is contained in any cycle we could decrease the cost by pushing more and more flow around this cycle until the arcs 
% capacity is at its limits.
% 
% This master thesis will deal with the problem of finding a network flow with no directed cycles (acyclic flow), which at 
% the same time maximizes the amount of flow on a specified arc $e$ of the network. We will discuss the complexity, an 
% exact algorithm based on a mixed integer program with separation of inequalities that forbid cycles and also a heuristic 
% approach that yields results much faster (but not optimal).%TODO am ende genauer schreiben was wirklich gemacht wurde
% 
% \subsection{The gas flow problem}
% Although it is mainly the motivation, not really the topic of this thesis, we want to briefly introduce the gas 
% transport problem. For more a detailed description we refer to LINK %TODO link auf ein entsprechendes ZIB-paper!!! 
